---
title: "TSA - ULSD"
author: "Nathaniel McNalley, Hongtian Fu, Taylor Armbruster, and Fashuai Li"
date: "`r Sys.Date()`"
output: html_document
resource_files:
- TSA_template.Rmd
- TSA_template.Rmd
---

<style type="text/css"> body, td {font-size: 12px;} code.r{font-size: 10px;} pre {font-size: 10px} </style>

```{r, echo = F, warning=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = F,warning = F,message = F, fig.width = 4, fig.height = 3,fig.align = "center",tidy = FALSE, strip.white = TRUE)

library(RTL)
library(plotly)
library(dplyr)
library(ggplot2)
library(tidyr)
library(purrr)
library(scales)

#Morningstar API Login
iuser = "morningstar.ctrd@ualberta.ca"
ipassword = "oRakvwtPVg"
EIAkey = "2a8946fed36919b63af3ff2f4bcaf83a"
```
### Summary
The main futures contracts used to trade ULSD will be the NY Harbor ULSD Futures (ticker HO). These contracts are standardized to 42,000 gallons (1,000 barrels). The ULSD delivery location is New York Harbor with options for delivery into the buyer's barge, tanker, or pipeline if they are able. The product must meet the specification of the Colonial Pipeline's Fungible Grade 62 for Ultra Low Sulfur Diesel and must be designated for sale in accordance with U.S. Environmental Protection Agency regulations. As of 2016, almost all petroleum-based diesel fuel available in the UK, mainland Europe, and North America is ultra low sulphur. 

NY Harbor ULSD Futures Handbook: https://www.cmegroup.com/content/dam/cmegroup/rulebook/NYMEX/1a/150.pdf  

ULSD is also commonly referred to as distillate and is the second-most consumed petroleum product in the United States. In addition to its use as a transportation fuel, distillate is also consumed for heating and power generation purposes. Distillate's use as a heating fuel drives the seasonal pattern of higher consumption during the winter months. Distillate consumption is affected by economic growth and weather conditions as well as vehicle efficiency and miles traveled of heavy-duty vehicles.

Components of the price you see at the pumps: The cost of crude oil purchased by refineries, refining costs, distribution and marketing costs, and state taxes all influence the price you see at the pumps.
Below are various charts that will be used as a quick dashboard for various monetization strategies. This includes flat price of ULSD and CL, as CL is a major component in ULSD production, the current crack spread between ULSD and CL, and the time spread between the first and second contract of ULSD.

### What's changed

Following a meeting with Professor Cote we determined that the March - April crack spread was too risky and difficult to model leading us to reverse our exposure in those months. 

There have been many instances of inventory draws over the past few months coupled with low storage from the Summer months. This lead us to believe that in the near term there would be limited likelihood that inventories would increase substantially prior to winter and that prices would continue to face upward pressure. In the last week we saw inventories build substantially in the U.S. leading to an increase in the days of supply available which contradicted our trading strategy that we entered into on November 1st. Due to these new developments we reversed our long time spread position and are now short the time spread given that inventories are now likely to build substantially going into Winter. 

### Current Exposure

Our current exposure is short the time spread between the two closest contracts of ULSD based on the belief that rising inventory in the near-term will lead to lower prices in the nearest contract compared to the 2nd contract due to higher demand later in the Winter months. We are short 250 December ULSD contracts and long 250 January ULSD contracts to monetize this belief. 

### Trades

October 22: Based on our initial market view that inventory levels going into the winter months would be insufficient to meet increased demand levels we chose to go long the March crack spread in anticipation of ULSD price appreciation relative to CL. We were long 15 contracts of ULSD and short 15 contracts of CL. 

October 29: When running regression analysis between the time spread of the two closest forward contracts and the days of inventory of ULSD we found that the current c1c2 time spread is on the higher side of the confidence interval. Given our market view that supply in the short-term will be inelastic relative to growing demand we anticipate continued storage draws leading to lower days of supply and a growing time spread. With days of supply falling to 30 or even less as the winter progresses the potential upside from monetizing the time spread could increase to above a dollar from its current 0.6 range in the market. To monetize this strategy we will be long 20 December 2021 and short 20 January 2022 ULSD contracts. The regression used for determining the reasonable range of the time spread based on historical data can be found in the Appendix. 

November 5: Reversed our Long Crack Spread position due to high volatile and difficulty in building and maintaining a sustainable model.

November 8: Storage builds appear to be increasing going into the winter months so we have decided to reverse our long December - January time spread strategy and instead short the time spread as inventories are likely to build over the next few weeks. This will lead to higher days of supply in the near-term and will decrease any upside that was available from being long the time spread. We have instead decided to be short 25 December ULSD contracts and long 25 January ULSD contracts in an attempt to benefit from the rising inventory levels in the near-term and the downward pressure it should put on prices.

November 11: Increased our position size for the short time spread strategy which we entered on November 9th. We are short a total of 250 December ULSD contracts and long 250 January ULSD contracts. 

### Supply Demand Balance

```{r}
#Implied storage builds and draws

data("tickers_eia")
sd <- tickers_eia %>% dplyr::filter(sd_category == "dist", category != "stocks") %>% 
dplyr::as_tibble()

eia_df <- tibble::tribble(~ticker, ~name) %>% 
  add_row(ticker = sd$tick.eia[1:nrow(sd)], name = sd$tick.r[1:nrow(sd)]) %>% 
  dplyr::mutate(key = EIAkey) %>% 
  dplyr::mutate(df = purrr::pmap(list(ticker, key, name), .f = RTL::eia2tidy)) %>% 
  dplyr::select(df) %>% tidyr::unnest(df)

data2 <- eia_df %>% 
  tidyr::pivot_wider(date, names_from = series, values_from = value) %>% 
  dplyr::rename(Demand = eia.mdist.demand, Imports = eia.mdist.imports, Supply = eia.mdist.supply, Exports = eia.mdist.exports) %>%   dplyr::arrange(date) %>% 
  dplyr::filter(date >= "2018-01-01") %>% 
  tidyr::drop_na() %>% 
  dplyr::mutate(Storage = Supply + Imports - Demand - Exports)

#Graphical Transformation Since 2018
fig.title = "ULSD SD Balance with Implied Builds/Draws Since 2018"
p1 <- data2 %>% 
  tidyr::pivot_longer(-date, names_to = "series", values_to = "value") %>% 
  ggplot(aes(x = date, y = value, color = series)) + 
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = fig.title, y = "kbd", x = "")
p1 %>% plotly::ggplotly(width = 700, height = 400,)

#Graphical Transformation Since 2021
fig.title = "ULSD SD Balance with Implied Builds/Draws Since 2021"
p2 <- data2 %>% 
  dplyr::filter( date>= "2021-01-01") %>% 
  tidyr::pivot_longer(-date, names_to = "series", values_to = "value") %>% 
  ggplot(aes(x = date, y = value, color = series)) + 
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = fig.title, y = "kbd", x = "")
p2 %>% plotly::ggplotly(width = 700, height = 400,)
```

Production: Most of the diesel fuel produced and consumed in the US is refined from crude oil at petroleum refineries within the US. These refineries produce an average of 11 to 12 gallons of diesel fuel for every 42 gallon barrel of crude. Production decisions by refineries are also effected by the spread between diesel and gasoline prices as when the spread widens refineries increase the amount of diesel that they produce relative to gasoline. The standard crack-spread at refineries is generally 3-2-1, meaning that for 3 barrels of crude that 2 will be refined into gasoline and 1 will be refined into ULSD. In our SD balance Supply represents U.S. domestic production of ULSD. ULSD production is vitally important to the functioning of the U.S. economy as it is the dominant fuel for freight transportation and heavy machinery. 

In the short term we believe that supply is relatively inelastic with production in the U.S. accounting for roughly 4.8 million barrels fo ULSD a day. In the longer term production is more elastic allowing refineries to increase the amount of ULSD that they produce if market conditions are accommodative. This would include higher ULSD prices from heightened demand. From our SD balance we can see the weekly changes in U.S. ULSD supplies. 

Consumption: More than 8 million hones in New England and the Central Atlantic Region of the US lack access to natural gas so they use Heating Oil (ULSD). This makes one of ULSD’s end markets consumers those who require ULSD for their home heating needs. Due to this there is some seasonality in pricing during the cold winter months. Residential heating oil consumption peaked in the 1970's and declined nearly every year since. 81%  of residential heating use is concentrated in the US Northeast including New York, Pennsylvania, and New England. Diesel is mainly used as a transport fuel and consumption is impacted by the underlying economic conditions, fuel efficiency, and miles traveled by heavy duty vehicles. In 2020, diesel fuel consumption was about 44.61 billion gallons which averages to about 3 million barrels per day. PADD 3 has been a net exported of diesel since 2001 which is logical given the number of refineries located in the US Gulf Coast and the fuel needs across the US. In our SD balance Demand represents the consumption of ULSD. Demand is important as it gives us a sense of the weekly needs of the U.S. for heating and transportation. 

In the short term we see demand increasing going into the holiday season as supply chain issues will require higher volumes of transport of freight across the US. Additionally, we are headed into the colder months of the year which will contribute to excess demand for home heating needs in the U.S. Northeast. In the longer term we believe that demand will stabilize following the winter season and after much of the U.S. supply chain congestion has been remedied. From our SD balance we are able to see weekly changes in the demand for ULSD. 

Net Exports: The U.S. is a net exporter of ULSD meaning that they export far more ULSD than they import. This is important as domestic production is being directed to producing ULSD which will be sent internationally instead of for meeting domestic demand and building inventories. Our SD balance shows both imports and exports for the U.S. meaning that we can analyze them separately. International trade of ULSD is very important as ULSD is being redirected to outside markets instead of being used domestically and is materially important for analyzing inventory builds and draws.

In the short term we believe that exports will decrease relative to imports to attempt and meet domestic needs during the winter months. In the longer term this pattern is likely to continue as exports will be rerouted to meet domestic demand needs. 

Storage: Inventories of ULSD are vitally important in the U.S. as higher demand in the winter months due to home heating needs can cause inventory draws. Consistent draws can lead to lower inventories and will eventually put upward pressure on prices to attract ULSD to that market. Inventory draws and builds are shown in the SD balance as Storage. Generally, during the summer months and approaching the winter months the U.S. will build up inventories of ULSD in preparation for higher demand in the winter. 

In 2021 it appears that due to high demand and exports relative to supply inventories were not built up. In the short term we expect that the lack of inventory builds will lead to higher prices for ULSD in the near term, especially in the later winter months closer to March. In the longer term as supply chain congestion is resolved and demand lessens for home heating we expect inventories to begin to build as they would typically. 

Transportation: Most diesel fuel moves by pipeline from refineries and ports to terminals near major consuming areas. Barges and trains also move diesel to terminals. Trucks transport the diesel fuel from the terminals to retail service stations and to large volume consumers such as fleet operators. Most of the U.S. refining capacity can be found in PADD 3 around the Gulf Coast and as such PADD 3 is a net exporter of ULSD. A large amount of demand in the colder winter months is located in PADD 1 and due to their low refining capacity they require imports of ULSD from refineries concentrated in PADD 3.

### Market View

What we see currently in the ULSD market is that refinery output for the next few quarters is constrained with any increases in domestic US supply being the result of a decrease in net exports in the short-term. In this sense domestic supply is only expected to increase by ~300 thousand barrels a day going into the winter season. Currently the market is in backwardation, meaning that futures prices are decreasing with increases in delivery timing. The forward curve can be found in Appendix 1.

With supply accounted for we need to make some assumptions about demand in the short and long term. In the short term we expect to see higher demand as a result of the cold winter months and home heating requirements, along with a continue push to solve supply chain issues across the US which would require an increase in freight transport during the holiday shopping season. Each of these factors leads us to believe that demand for ULSD in the next few quarters will be strong and outstrip US supply in the near-term. 

Going into the winter months we would expect to see inventories increasing but with the prolonged downward trend it appears that inventory builds will solely be driven by changes in net exports as supply from refineries appears to be inelastic. If demand in the US market during the winter months increases more than the change in net exports we can expect to see increases in inventory draws and decreases in days of distillate supply. 

US distillate oil days of supply has been on a downward trend since late January 2021 with less than 36 days of supply since August supporting the idea that inventories have not been able to build much going into winter.

Both historical ULSD inventories and days of supply have been charted to get a sense of current storage levels and how they have been trending. ULSD inventories are forecasted by EIA so there are data points as far as the end of 2022. Days of supply are based on realized values and are updated weekly and total inventories are reported monthly. Both charts can be found in Appendix 1. A chart was also created to illustrate the relationship between flat price, time spreads, and crack spreads in relation to days of supply and can be found in Appendix 4. 

### Desired Exposure

As we believe that demand will outstrip supply during the winter season we will want to be long during this period. With increased inelastic demand due to transport needs and higher demand due to heating oil needs in the winter coupled with limited inventory builds during the summer we should see higher inventory draws leading to lower days of supply. The anticipated downward trend in days of supply during the winter will likely be the catalyst for higher ULSD prices. To monetize our bullish market call we could use the following strategies:

Flat Price: Long contracts of ULSD during the winter months to capitalize on higher demand and sustained inventory draws. The chart for the flat price of the closest ULSD contract can be found in Appendix 2.

Time Spread: Exposure to the difference between ULSD contracts with different delivery dates. Would involve taking a long position in nearer term contracts and a short position in ULSD contracts past March. The chart for the time spread of the closest two ULSD contracts and the March - April contracts can be found in Appendix 2. 

Crack Spread: Would involve exposure to CL prices which is a major determinant in the price of ULSD. Establishing a long position in the crack spread with the belief that ULSD prices will appreciate relative to CL prices during the winter months. Most likely exposure taken in March as inventories are generally their lowest at the end of the winter and unexpected cold weather can lead to major increases in ULSD prices relative to CL. The March crack spread chart can be found in Appendix 2. 

### Monetization Strategies

The three main monetization strategies mentioned in the desired exposure section above were as follows: long the flat price, long the time spread, and long the crack spread. To get a sense of the historical performance of these strategies the historical P/L and value charts for these strategies are included below. 
```{r}
tradeprocess <- RTL::getPrices(feed="CME_NymexFutures_EOD",contracts = c("@CL22H","@HO22J","@HO22H"),
                          from = "2012-01-01",iuser = iuser, ipassword = ipassword) %>% stats::na.omit()

df.long <- tradeprocess %>% tidyr::pivot_longer(cols = -date,
   names_to = "series",
  values_to = "value") %>%
  dplyr::mutate(value = case_when(
    grepl("HO", series) ~ value * 42,
    TRUE ~ value
  ))

# Plot into two panels using facet_grid
fp <- df.long %>%
  dplyr::filter(series == "HO22H") %>%
  dplyr::mutate(PL = value - dplyr::lag(value)) %>%
  stats::na.omit() %>%
  dplyr::select(-series) %>%
  tidyr::pivot_longer(cols = -date,
                      names_to = "series",
                      values_to = "value")

# Plot into two panels using facet_grid
fp1 <- fp %>%
  ggplot(aes(x = date, y = value)) + geom_bar(col = "blue", stat = "identity") +
  facet_grid(series ~ ., scales = "free_y") +
  labs(
    title = "Strategy::March Flat Price",
    subtitle = "Price Evolution and Risk/Reward",
    caption = "Source: Morningstar/CME",
    x = "",
    y = "$ per bbl"
  )
fp1 %>% plotly::ggplotly(width = 700, height = 400,)
```

```{r}
cx <- df.long %>%
  dplyr::filter(series %in% c("HO22H", "CL22H")) %>%
  tidyr::pivot_wider(names_from = series, values_from = value) %>% 
  dplyr::transmute(
    date = date,
    value = HO22H - CL22H,
    PL = value - dplyr::lag(value)
  ) %>%
  stats::na.omit() %>%
  tidyr::pivot_longer(cols = -date,
                      names_to = "series",
                      values_to = "value")

# Plot into two panels using facet_grid
cx1 <- cx %>%
  ggplot(aes(x = date, y = value)) + geom_bar(col = "blue", stat = "identity") +
  facet_grid(series ~ ., scales = "free_y") +
  labs(
    title = "Strategy::March Crack Spread",
    subtitle = "Price Evolution and Risk/Reward",
    caption = "Source: Morningstar/CME",
    x = "",
    y = "$ per bbl"
  )
cx1 %>% plotly::ggplotly(width = 700, height = 400,)
```

```{r}
struct <- df.long %>%
  dplyr::filter(series %in% c("HO22J", "HO22H")) %>%
  tidyr::pivot_wider(names_from = series, values_from = value) %>% 
  dplyr::transmute(
    date = date,
    value = HO22H - HO22J,
    PL = value - dplyr::lag(value)
  ) %>%
  stats::na.omit() %>%
  tidyr::pivot_longer(cols = -date,
                      names_to = "series",
                      values_to = "value")

# Plot into two panels using facet_grid
Str <- struct %>%
  ggplot(aes(x = date, y = value)) + geom_bar(col = "blue", stat = "identity") +
  facet_grid(series ~ ., scales = "free_y") +
  labs(
    title = "Strategy::March - April Time Spread",
    subtitle = "Price Evolution and Risk/Reward",
    caption = "Source: Morningstar/CME",
    x = "",
    y = "$ per bbl"
  )
Str %>% plotly::ggplotly(width = 700, height = 400,)
```

### Risk Appetite
Distributions were also created to get a sense of the volatility of each of the strategies and for comparative purposes. The chart below compares the payoff distributions of each of the strategies. 

```{r}
# Using row bind (rbind) to concatenate long dfs 
res <-
  rbind(
    fp %>% dplyr::filter(series != "value") %>% dplyr::mutate(series = "Flat Price"),
    cx %>% dplyr::filter(series != "value") %>% dplyr::mutate(series = "Crack Spread"),
    struct %>% dplyr::filter(series != "value") %>% dplyr::mutate(series = "Time Spread")
  )

Res <- res %>% ggplot(aes(x = value)) +
  stat_density(mapping = (aes(y = ..scaled..)),
               col = "blue",
               fill = "blue") +
  facet_grid(series ~ .) +
  labs(
    title = "Strategy PL Comparison",
    caption = "Source: Morningstar",
    x = "",
    y = "$ per bbl"
  )
Res %>% plotly::ggplotly(width = 700, height = 400,)
```
Based on the distributions of each strategy it appears that the flat price strategy has the highest risk and the March - April 2022 time spread strategy has the lowest risk. The crack spread has much more moderate risk compared to the flat price strategy but has a much wider distribution than the time spread strategy. The crack spread strategy would also expose us to changes in the price of CL which may be beneficial if there are increases in CL production leading to moderate price declines. 

We feel that the distribution and risk of the flat price strategy is too high for making trades and that the other two strategies have much more reasonable risk distributions. At this time we are comfortable starting with a higher risk, higher reward crack spread strategy and we will likely include the lower risk time spread strategy as a later trade. 

```{r}
library(tibbletime)
# Using RTL and Morningstar to retrieve contract data
df <-  RTL::getPrices(
  feed = "CME_NymexFutures_EOD_continuous",
  contracts = paste0("HO_", sprintf('%0.3d', 1:36), "_Month"),
  from = "2015-01-01",
  iuser = iuser,
  ipassword = ipassword
) %>%
  dplyr::filter(!date %in% c(as.Date("2020-04-20")))

df <- df %>%
  tidyr::pivot_longer(-date, names_to = "series", values_to = "value") %>%
  dplyr::mutate(series = stringr::str_replace_all(series, c("_0" = "", "_Month" = ""))) %>%
  dplyr::group_by(series) %>%
  dplyr::mutate(
    ret_abs = value - dplyr::lag(value),
    ret_rel = (value / dplyr::lag(value)) - 1,
    ret_log = log(value / dplyr::lag(value))
  ) %>%
  stats::na.omit()
## Pivot prices into a wide data frame
prices <- df %>% dplyr::select(date, series, value) %>%
  tidyr::pivot_wider(values_from = value, names_from = series)
```

```{r}
## Pivot returns into a wide data frame
ret_abs <- df %>% dplyr::select(date, series, ret_abs) %>%
  tidyr::pivot_wider(values_from = ret_abs, names_from = series)
ret_log <- df %>% dplyr::select(date, series, ret_log) %>%
  tidyr::pivot_wider(values_from = ret_log, names_from = series)
```

```{r}
# Roll Adjustment
library(moments)
ret.log.adj <-
  RTL::rolladjust(
    x = ret_log,
    commodityname = c("cmeulsd"),
    rolltype = c("Last.Trade")
  ) %>%
  tidyr::pivot_longer(-date, names_to = "series", values_to = "value")

x <- ret.log.adj %>%
  stats::na.omit() %>%
  dplyr::group_by(series) %>%
  dplyr::summarize(
    mean = mean(value),
    stddev = sd(value),
    kurtosis = moments::kurtosis(value),
    skewness = moments::skewness(value),
    .groups = "keep"
  )
```

Volatility in ULSD was found to decrease when the time to delivery increases. This makes sense as prices are more susceptible to changes in supply and demand in the near-term and less susceptible in the longer-term due to increased elasticity over time. The three charts below illustrate this relationship between delivery timing and volatility. 

```{r}
fig.title = "Risk versus Time to Delivery"

sd.simple <- ret.log.adj %>%
  group_by(series) %>%
  dplyr::summarize(stddev = sd(value) * sqrt(252), .groups = "keep")

SD.Simple <- sd.simple %>% ggplot(aes(x = series, y = stddev)) + geom_point(col = "blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(labels = percent) +
  labs(
    title = fig.title,
    subtitle = "notice the volatility cone shape: risk is a function of maturity",
    x = "Contracts",
    y = "Annualized Percentage Volatility"
  )
SD.Simple %>% plotly::ggplotly(width = 700, height = 400,)
```

```{r}
fig.title = "Box Plot for Risk and Time to Delivery"
 ret.log.adj %>% plot_ly(
  y = ~ value,
  color = ~ series,
  type = "box",
  colors = c("blue", "orange"),
  width = 700,
  height = 400
)  %>%
  plotly::layout(
    title = list(text = fig.title, x = 0),
    xaxis = list(title = ""),
    yaxis = list(
      title = "Daily Returns",
      separators = '.,',
      tickformat = ".0%"
    )
  )
```

```{r}
library(tibbletime)
# Define the rolling function using rollify()
sd20 <- tibbletime::rollify(stats::sd, window = 20)
# We use long data frames in large data sets
ret.roll <- tibbletime::as_tbl_time(ret.log.adj, index = date) %>%
  dplyr::group_by(series) %>%
  dplyr::mutate(sd20 = sd20(value) * sqrt(252)) %>%
  dplyr::ungroup() %>% stats::na.omit()
# Since series is numeric, we use as.factor to have discrete scale
ret.roll %>% dplyr::filter(series %in% c("HO01", "HO06", "HO12", "HO36")) %>%
  plotly::plot_ly(
    x = ~ date,
    y = ~ sd20,
    name = ~ series,
    color = ~ series,
    width = 700,
  height = 400
  ) %>%
  plotly::add_lines() %>%
  plotly::layout(
    title = list(text = "Heteroskedasticity", x = 0),
    xaxis = list(title = ""),
    yaxis = list(title = "Annualized Volality", tickformat = ".0%")
  )
```

Correlations between ULSD were also tested to get a sense of how much a change in the front contract can have on a contract one year and three years from now. The result is consistent with the findings of the previous volatility charts as closer contracts will have a higher correlation with one another when compared to two contracts with a wide gap in delivery timings. In the long term, supply and demand are more elastic meaning that consumers and producers can change their behaviors leading to more moderate price changes farther down the curve. In the near-term demand and supply are more inelastic so price changes are more volatile in response to immediate inventory levels and demand swings. Correlations between ULSD contracts are larger when the contract delivery dates are closer together and smaller when delivery times are farther apart. This means that time has a mitigating effect on the near-term price volatility of ULSD as closer contracts will move in line with one another but farther out contracts will not experience price volatility to the same extent. The chart below illustrates this relationship. 
```{r}
ret.abs.adj <-
  RTL::rolladjust(
    x = ret_abs,
    commodityname = c("cmeulsd"),
    rolltype = c("Last.Trade")
  )
```

```{r}
fig.title = "Rolling Correlations"
# Define Rolling Cor from stats::cor() with n=20
# .x and .y and the two generic variables over which we compute return correlation
cor_roll <- tibbletime::rollify( ~ cor(.x, .y), window = 20)

# Data Set in Tibbletime
# Requires time based objects
# Connot cope with non-alphanumeric column names when calling variables in a function

df.corr <- ret.abs.adj %>%
  dplyr::select(date, HO01, HO03, HO12, HO36) %>%
  tibbletime::as_tbl_time(index = date) %>%
  stats::na.omit()

# Applying
df.corr <- df.corr %>%
  dplyr::mutate(
    rho_1_3 = cor_roll(.x = HO01, .y = HO03),
    rho_1_12 = cor_roll(.x = HO01, .y = HO12),
    rho_1_36 = cor_roll(.x = HO01, .y = HO36)
  ) %>%
  stats::na.omit()

# Charting With Minimal Edits if adding more correlations
 df.corr %>% dplyr::select(date, contains("rho")) %>%
  tidyr::pivot_longer(cols = -date,
                      names_to = "series",
                      values_to = "value") %>%
  plotly::plot_ly(
    x = ~ date,
    y = ~ value,
    name = ~ series,
    color = ~ series,
    width = 700,
  height = 400
  ) %>%
  plotly::add_lines() %>%
  plotly::layout(
    title = list(text = fig.title, x = 0),
    xaxis = list(title = ""),
    yaxis = list(title = fig.title, tickformat = ".2f")
  )
```

### Execution

October 22: The first trade that we will be executing will be a crack spread strategy for March in which we are long ULSD and short CL. From our supply demand balance production of ULSD is quite inelastic in the short term and that and increases in domestic U.S. supply will have to be the result of a decrease in exports. Only slight and temporary decreases in exports are expected prior to the winter season so it is not likely that domestic inventories will build up enough to meet demand into the late winter. There was no build up of inventories in the Summer as would be typically suspected resulting in lower inventories going into the Fall with moderate draws of inventory already being realized leading to lower days of supply. We believe that higher demand during the winter season will be fueled by residential heating oil demand and freight transport needs over the holiday season with a continued effort to resolve U.S. supply chain bottlenecks leading to higher than average demand for ULSD. This will leads to draws on inventory which we anticipate will cause the crack spread between ULSD and CL to widen near to February and March as inventories have historically bottomed in those months. The relationship between days of supply and the crack spread appeared to be inverse in historical periods so based on our belief that we will see moderately lower days of inventory in the late winter we expect to profit off the widening of the crack spread. EIA forecasts show that inventories will be at their lowest point for the winter season in March and April which is why we decided to long the March crack spread. 

This will be a longer term strategy that we would expect to pay off closer to the end of March depending on the severity of the winter weather at that time. We do expect to profit off of this trade as we move closer to March and as days of supply decrease after the initial builds due to the decline in ULSD exports. 

Currently the crack spread for March 2022 is ~24 dollars per barrel which we believe is quite low given our expectations of continued inventory draws and lower days of supply during the Winter making the current spread an attractive entry point. Comparable days of supply to today have realized historical crack spreads between ~26 to ~40 dollars depending on the severity of the winter making this trade quite attractive. Trade exit would take place at a price in this range unless market conditions change materially. This trade will involve being long 15 contracts of ULSD and short 15 contracts of CL.

October 29: When running regression analysis between the time spread of the two closest forward contracts and the days of inventory of ULSD we found that the current c1c2 time spread is on the higher side of the confidence interval. Given our market view that supply in the short-term will be inelastic relative to growing demand we anticipate continued storage draws leading to lower days of supply and a growing time spread. With days of supply falling to 30 or even less as the winter progresses the potential upside from monetizing the time spread could increase to roughly 0.99 from its current 0.6 range in the market. To monetize this strategy we will be long 20 December 2021 and short 20 January 2022 ULSD contracts. We will exit this position if the market changes dramatically, if storage builds unexpectedly increase, or if the spread reaches the higher end of our confidence interval $0.99. (The regression and prediction values can be found in Appendix 5. The monetization and risk distribution for the front time spread strategy can be found in Appendix 3. 
November 5: We exited our long March crack spread position due to recommendations that it is quite volatile and difficult to build and maintain a reliable model.

November 8: Storage builds have suddenly increased leading us to reevaluate our long time spread position. With new storage builds and increasing days of supply we decided to reverse our long December - January time spread and instead short the time spread as inventories are likely to build over the next few weeks leading decreases in price and our potential upside.

November 11: We decided to increase our risk with regard to the short time spread that we initiated on November 8th resulting in a total of 250 contracts. 

### Profit and Loss

Below are the profit and loss graphs for the two strategies that we deployed over the trading game. Neither strategies generated positive returns at this time but there were instances where we had paper profits for the time spread strategy. There is a large decline in profitability in the time spread strategy on November 11th as we entered significantly more contracts.
Generally we learned that we needed to deploy much more risk than what we had been over the game.
```{r}
#Crack Spread P/L
tradedate1 <- as.Date("2021-10-22")
tradedate2 <- as.Date("2021-11-05")

ULSD_H <- getPrices(feed = "CME_NymexFutures_EOD",
                    contracts = "@HO22H",
                    from = tradedate1, iuser = iuser, ipassword = ipassword) %>% 
          dplyr::filter(date <= tradedate2)
CL_H <- getPrices(feed = "CME_NymexFutures_EOD",
                    contracts = "@CL22H",
                    from = tradedate1, iuser = iuser, ipassword = ipassword) %>% 
          dplyr::filter(date <= tradedate2)

#P and L on a single contract basis
trade1 <- mutate(CL_H, shortat= 79.99, ULSD_H, longat = 2.487, LongPL = (ULSD_H$HO22H - longat)*42, ShortPL = (shortat - CL_H$CL22H), TotalPL =LongPL+ShortPL)

#Total P and L on total trade
Trade1 <- ggplot(trade1, aes(x=date,y=TotalPL*15)) +geom_line() +ggtitle("Long March Crackspread P/L") + ylab("profit/loss")+ scale_y_continuous(labels = comma)

Trade1 %>% plotly::ggplotly(width = 700, height = 400,)
```


```{r}
tradedate1 <- as.Date("2021-10-29")
tradedate2 <- as.Date("2021-11-08")
tradedate3 <- as.Date("2021-11-08")
tradedate4 <- as.Date("2021-11-11")

HOLong1 <- getPrices(feed = "CME_NymexFutures_EOD",
                    contracts = "@HO21Z",
                    from = tradedate1, iuser = iuser, ipassword = ipassword) %>% 
          dplyr::filter(date <= tradedate2)

HOShort1 <- getPrices(feed = "CME_NymexFutures_EOD",
                    contracts = "@HO22F",
                    from = tradedate1, iuser = iuser, ipassword = ipassword) %>% 
          dplyr::filter(date <= tradedate2)

HOLong2 <- getPrices(feed = "CME_NymexFutures_EOD",
                    contracts = "@HO22F",
                    from = tradedate3, iuser = iuser, ipassword = ipassword)

HOShort2 <- getPrices(feed = "CME_NymexFutures_EOD",
                    contracts = "@HO21Z",
                    from = tradedate3, iuser = iuser, ipassword = ipassword) 

HOLong3 <- getPrices(feed = "CME_NymexFutures_EOD",
                    contracts = "@HO22F",
                    from = tradedate4, iuser = iuser, ipassword = ipassword)

HOShort3 <- getPrices(feed = "CME_NymexFutures_EOD",
                    contracts = "@HO21Z",
                    from = tradedate4, iuser = iuser, ipassword = ipassword) 

date_range <- getPrices(feed = "CME_NymexFutures_EOD",
                    contracts = "@HO21Z",
                    from = tradedate1, iuser = iuser, ipassword = ipassword) %>% 
                dplyr::select(date)


#P and L for Initial Long Time Spread Trade
trade2 <- mutate(HOLong1, longat= 2.5014, HOShort1, shortat = 2.4865, pl1 = ((HOLong1$HO21Z-longat)*42)*20, pl2 = ((shortat-HOShort1$HO22F)*42)*20, totalpl2=pl1+pl2) 

#P and L for Initial Short Time Spread Trade
trade3 <- mutate(HOLong2, longat= 2.501, HOShort2, shortat = 2.5153, pl1 = ((HOLong2$HO22F-longat)*42)*25, pl2 = ((shortat-HOShort2$HO21Z)*42)*25, totalpl3=pl1+pl2) 

#P and L for Increased Risk on Time Spread Trade
trade4 <- mutate(HOLong3, longat= 2.391, HOShort3, shortat = 2.3991, pl1 = ((HOLong3$HO22F-longat)*42)*225, pl2 = ((shortat-HOShort3$HO21Z)*42)*225, totalpl4=pl1+pl2) 

trade2 <- dplyr::select(trade2, date, totalpl2)
trade3 <- dplyr::select(trade3, date, totalpl3)
trade4 <- dplyr::select(trade4, date, totalpl4)

df <- dplyr::left_join(date_range,trade4) 
df <- dplyr::left_join(df, trade3)
df <- dplyr::left_join(df,trade2)

df <- df %>% 
  dplyr::group_by(date) 
 df[is.na(df)] <- 0
df <- df %>% dplyr::summarise(totalpl = sum(totalpl2+totalpl3+totalpl4))

Trade2 <- ggplot(df, aes(x=date,y=totalpl)) +geom_line() + scale_y_continuous(labels = comma) +ggtitle("Time Spread Strategy P/L") + ylab("profit/loss")
Trade2 %>% plotly::ggplotly(width = 700, height = 400,)
```


### Lessons Learned

It is difficult but essential to keep on top of the market and your positions at all times. The market can change quite quickly and if your trades are stale you can end up getting hurt.

Questions are an ever present part of learning and with an environment that is so large and quickly changing it is essential to ask questions. 

### Appendicies

### Appendix 1

```{r}
#Pulling the Forward Curve for HO
fig.title = "Heating Oil Futures Forward Curve"
HO_Curve <- RTL::getCurve(
  feed = "Crb_Futures_Price_Volume_And_Open_Interest",
  contract = "HO",
  date = "2021-10-29",
  iuser = iuser,
  ipassword = ipassword
) %>% 
  dplyr::select(c("contract", "Close")) %>% 
  dplyr::mutate(Close = Close * 42)
HO_Curve %>% 
  plot_ly(
     width = 700, height = 400,
    x = ~contract,
    y = ~Close,
    type = "scatter",
    mode = "markers",
    name = "Forward Curve") %>% 
  plotly::layout(
    title = list(text = fig.title, x = 0),
    xaxis = list(
      title = "Contract",
      categoryorder = "array",
      categoryarray = ~ contract
    ),
    yaxis = list(title = "$ per bbl")
  )
```

```{r}
#Pulling Distillate inventories
eia_df <- 
  tibble::tribble(
  ~ticker, ~param, ~type,
  "STEO.DFPSPUS.M", "Inventory", "Source"
  )
data <- eia_df %>%
  dplyr::mutate(key = EIAkey) %>%
  dplyr::mutate(data = purrr::map2(ticker, key, RTL::eia2tidy)) %>%
  dplyr::select(param, type, data) %>% unnest() %>% 
  dplyr::filter(date>="2018-01-01")

p1 <-
  data %>%
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle('Distillate Fuel Oil U.S. Total Inventory, Monthly')
p1 %>% plotly::ggplotly(width = 700, height = 400,)
```

```{r}
#pulling in days of supply of distillate
eia_df <- 
  tibble::tribble(
  ~ticker, ~param, ~type,
  "PET.W_EPD0_VSD_NUS_DAYS.W", "Value", "Source"
  )
days_supply <- eia_df %>%
  dplyr::mutate(key = EIAkey) %>%
  dplyr::mutate(data = purrr::map2(ticker, key, RTL::eia2tidy)) %>%
  dplyr::select(param, type, data) %>% unnest(cols = c(data)) %>% 
  dplyr::select(date, series, value)

p1 <- days_supply %>%
  dplyr::filter(date >= "2018-01-01") %>% 
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle('Distillate Fuel Oil Days of Supply U.S, Weekly')
p1 %>% plotly::ggplotly(width = 700, height = 400,)
```

### Appendix 2

```{r}
#Extracting the March and April Contracts
df <- RTL::getPrices(
  feed = "CME_NymexFutures_EOD",
  contracts = c("@HO22H","@HO22J","@CL22H","@CL22J"),
  from = "2010-01-01",
  iuser = iuser,
  ipassword = ipassword) 
#Extracting the 1st and 2nd Contracts
df2 <- RTL::getPrices(
    feed = "CME_NymexFutures_EOD_continuous",
    contracts = c("HO_001_Month","HO_002_Month","CL_001_Month","CL_002_Month"),
    from = "2010-01-01",
    iuser = iuser,
    ipassword = ipassword)
#Joining the Data Frames, converting to barrels, and calculating spreads
HOdf <- dplyr::inner_join(df,df2) %>% 
    dplyr::arrange(dplyr::desc(date)) %>% 
  stats::na.omit()%>% 
  dplyr::transmute(date = date,
                   HO_01 = HO_001_Month * 42,
                   HO_02 = HO_002_Month * 42,
                   CL_01 = CL_001_Month,
                   CL_02 = CL_002_Month,
                   HO22H = HO22H * 42,
                   HO22J = HO22J * 42,
                   CL22H = CL22H,
                   CL22J = CL22J,
                   HO_Current_Crack = HO_01 - CL_01,
                   HO_c1c2_Spd = HO_01 - HO_02,
                   HO_Apr22_Mar22_Spd = HO22H - HO22J,
                   HO_Crack_Spd = HO22H - CL22H) %>% 
#Converting from wide to long
  tidyr::pivot_longer(-date, values_to = "Price", names_to = "Contracts")
```

```{r}
#Plotting flat price for 1st HO Contract
HOdf %>% 
  dplyr::filter(stringr::str_detect(Contracts, "HO_01")) %>% 
  dplyr::filter(date >= "2021-01-01") %>% 
  plotly::plot_ly(
     width = 700, height = 400,
    x = ~date,
    y = ~Price,
    name = ~ Contracts,
    type = 'scatter',
    mode = 'lines'
  ) %>% 
  plotly::layout(
    title = "Flat Price of USLD", x = 0,
  xaxis = list(title = "Date"),
  yaxis = list(
    title = "$ per Barrel")
  )
```

```{r}
#Time spread 1st and 2nd contracts for ULSD
HOdf %>% 
  dplyr::filter(stringr::str_detect(Contracts, "c1c2")) %>% 
  plotly::plot_ly(
     width = 700, height = 400,
    x = ~date,
    y = ~Price,
    name = ~Contracts
  ) %>% 
  plotly::add_lines() %>% 
  plotly::layout(
    title = "Time Spread: HO_01 vs HO_02", x = 0,
  xaxis = list(title = "Date"),
  yaxis = list(
    title = "$ per Barrel")
  )
```

```{r}
#Time spread for March and April ULSD contracts 
HOdf %>% 
  dplyr::filter(stringr::str_detect(Contracts, "Apr")) %>% 
  plotly::plot_ly(
     width = 700, height = 400,
    x = ~date,
    y = ~Price,
    name = ~Contracts
  ) %>% 
  plotly::add_lines() %>% 
  plotly::layout(
    title = "Time Spread: March and April 2022", x = 0,
  xaxis = list(title = "Date"),
  yaxis = list(
    title = "$ per Barrel")
  )
```

```{r}
#Crack spread between ULSD and WTI
HOdf %>% 
  dplyr::filter(stringr::str_detect(Contracts, "Crack_S")) %>% 
  plotly::plot_ly(
     width = 700, height = 400,
    x = ~date,
    y = ~Price,
    name = ~Contracts
  ) %>% 
  plotly::add_lines() %>% 
  plotly::layout(
    title = "ULSD Crack Spread: WTI March 2022", x = 0,
  xaxis = list(title = "Date"),
  yaxis = list(
    title = "$ per Barrel")
  )
```

### Appendix 3
```{r}
#Front Time Spread Risk

Reg <- dplyr::filter(HOdf, Contracts == "HO_c1c2_Spd") %>% 
  dplyr::rename(value = "Price") %>% 
  dplyr::rename(series = "Contracts") %>% 
  tidyr::pivot_wider(names_from = series, values_from = value) %>% 
  stats::na.omit()

days_supply <- days_supply %>% 
  tidyr::pivot_wider(names_from = series, values_from = value) %>% 
  dplyr::rename(Days_Supply = PET.W_EPD0_VSD_NUS_DAYS.W)
 
c1c2_Reg <- dplyr::left_join(Reg, days_supply) %>% 
  dplyr::arrange(date) %>% 
  tidyr::fill(Days_Supply) %>% 
  stats::na.omit() %>% 
  dplyr::mutate(year = lubridate::year(date))
struct <- c1c2_Reg %>%
  dplyr::select(date, "HO_c1c2_Spd") %>% 
  tidyr::pivot_longer(-date, names_to = "series", values_to = "value")

# Plot into two panels using facet_grid
Str <- struct %>%
  ggplot(aes(x = date, y = value)) + geom_bar(col = "blue", stat = "identity") +
  facet_grid(series ~ ., scales = "free_y") +
  labs(
    title = "Strategy::Front Time Spread",
    subtitle = "Price Evolution and Risk/Reward",
    caption = "Source: Morningstar/CME",
    x = "",
    y = "$ per bbl"
  )
Str %>% plotly::ggplotly(width = 700, height = 400,)
```

```{r}
#Distribution
Res <- struct %>% ggplot(aes(x = value)) +
  stat_density(mapping = (aes(y = ..scaled..)),
               col = "blue",
               fill = "blue") +
  facet_grid(series ~ .) +
  labs(
    title = "Front Time Spread Distribution",
    caption = "Source: Morningstar",
    x = "",
    y = "$ per bbl"
  )
Res %>% plotly::ggplotly(width = 700, height = 400,)
```

### Appendix 4
```{r}
#Putting together days of supply and price strategies (crack spread, time spread, flat price)
#Extracting the 1st and 2nd Contracts
df2 <- RTL::getPrices(
    feed = "CME_NymexFutures_EOD_continuous",
    contracts = c("HO_001_Month","HO_002_Month","CL_001_Month"),
    from = "2010-01-01",
    iuser = iuser,
    ipassword = ipassword)
#Joining the Data Frames, converting to barrels, and calculating spreads
 Comp_df <-  df2 %>%  dplyr::arrange(dplyr::desc(date)) %>% 
  stats::na.omit()%>% 
  dplyr::transmute(date = date,
                   HO_01 = HO_001_Month * 42,
                   HO_02 = HO_002_Month * 42,
                   CL_01 = CL_001_Month,
                   Time_Spread = HO_01 - HO_02,
                   Crack_Spread = HO_01 - CL_01)
 
#Put together and converting from wide to long
Comp_df <- dplyr::inner_join(days_supply,Comp_df) %>% 
  dplyr::select(-HO_02 & -CL_01) %>% 
  dplyr::rename(Flat_Price = HO_01) %>% 
  tidyr::pivot_longer(-date, values_to = "value", names_to = "series")

fig.title = "Flat Price, Time Spread, and Crack Spread Compared to Days of Supply"
t <- Comp_df %>% 
  ggplot(aes(x = date, y = value, color = series)) + 
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = fig.title, y = "Value", x = "")
 t %>% plotly::ggplotly(width = 700, height = 400,)
```

### Appendix 5

```{r}
#Regressing Days of Supply and Time Spread.
Reg <- dplyr::filter(HOdf, Contracts == "HO_c1c2_Spd") %>% 
  dplyr::rename(value = "Price") %>% 
  dplyr::rename(series = "Contracts") %>% 
  tidyr::pivot_wider(names_from = series, values_from = value) %>% 
  stats::na.omit()

c1c2_Reg <- dplyr::left_join(Reg, days_supply) %>% 
  dplyr::arrange(date) %>% 
  tidyr::fill(Days_Supply) %>% 
  stats::na.omit() %>% 
  dplyr::mutate(year = lubridate::year(date))

c1c2_Reg %>% plotly::plot_ly(
  x = ~ Days_Supply ,
  y = ~ HO_c1c2_Spd,
  #name = ~ series,
  color = ~ year,
  height = 400,
  width = 700,
  type = "scatter",
  mode = "markers"
) %>%
  plotly::layout(
    title = list(text = "Front Spread Market Structure vs Days of Storage", x = 0),
    xaxis = list(title = "Days of Supply"),
    yaxis = list(title = "Time Spread ($)")
  )
```

```{r}
# Running the regression with the stats package
c1c2_Reg <- c1c2_Reg %>% dplyr::filter(abs(HO_c1c2_Spd) < 2)
fit <- lm(HO_c1c2_Spd ~ splines::bs(Days_Supply,knots = c(30,35), degree = 3, intercept = FALSE), data = c1c2_Reg)
summary(fit)

library(ggfortify)
autoplot(fit, size = 0.5)

c1c2_Graph <- c1c2_Reg %>% ggplot(aes(x = Days_Supply, y = HO_c1c2_Spd, col = year)) +
  geom_point() + 
    geom_smooth(
    method = "lm",
    formula = y ~ splines::bs(
      x,
      knots = c(30,35),
      degree = 3,
      intercept = T
    ),
    se = F,
    col = "red"
  ) 
c1c2_Graph %>% plotly::ggplotly(width = 700, height = 400,)

```

```{r}
#Prediction for November 1 
Days_Supply = data.frame(x = 30.4)
fit %>% stats::predict(newdata = Days_Supply,
                       interval = "prediction",
                       level = 0.99)
```
The confidence interval on the regression is much larger than we would have expected but is likely caused by outlier values realized in 2020 during the COVID-19 Pandemic. Using the regression results between the ULSD time spread and days of supply we can assess if the current time spread is reasonable based on the most recent reported days of supply. In this case the most recent values for days of supply is 30.4. When inputting this into the regression we find that the reasonable spread given a 99% confidence interval is between -0.80 and 0.99 while the current spread is roughly $0.64. The current market spread is within the confidence interval but is on the higher side of the distribution meaning that there is a smaller amount of upside potential based on the current days of supply and the regression. With the expectation that days of supply will continue to decrease we could reasonably see more significant upside from a long time spread strategy. 


### Appendix 6

```{r}
#Prediction at current days of supply
Days_Supply = data.frame(x = 32.3)
fit %>% stats::predict(newdata = Days_Supply,
                       interval = "prediction",
                       level = 0.99)
```

Using the regression results between the ULSD time spread and days of supply we can assess if the current time spread is reasonable based on the most recent reported days of supply. In this case the most recent values for days of supply is 32.3. When inputting this into the regression we find that the reasonable spread given a 99% confidence interval is between -0.916 and 0.868 while the current spread is roughly $0.57. The current market spread is within the confidence interval but with inventories building in the near-term going into winter it is likely that days of supply will increase further decreasing any potential upside.

### Heating Degree Day and Demand Analysis
```{r}
#Utilizing HDD and Forecast to Predict Inventory Levels and Potential Future Prices
#HDD used is for the Middle Atlantic region of the US including NY and Pennsylvania. Should be representative for the New England region as well.
fig.title = "ULSD Demand vs HDD"
x <-
  rbind(
    RTL::eia2tidy("STEO.ZWHDPUS.M", key = EIAkey, name = "HDD"),
    RTL::eia2tidy("STEO.DFTCPUS.M", key = EIAkey, name = "Demand")
  ) %>%
  dplyr::group_by(series) %>%
  dplyr::arrange(date) %>%
  dplyr::mutate(month = lubridate::month(date),
                change = value - dplyr::lag(value)) %>%
  stats::na.omit() %>%
  dplyr::filter(date > "2010-01-01",
                month %in% c(11, 12, 1, 2, 3)) %>%
  dplyr::select(date, series, change, month) %>%
  tidyr::pivot_wider(names_from = series, values_from = change)

HD <-  x %>%
  ggplot(aes(x = HDD, y = Demand, col = as.character(month))) +
  geom_point() +
  labs(title = fig.title)
HD %>% plotly::ggplotly(width = 700, height = 400,)

```
From the above scatterplot it does not appear that there is a clear linear relationship between ULSD demand and the winter months. Heating degree days tend to be concentrated together based on the month so the number of HDD in a month is fairly consistent each year. 

```{r}
# Warnings 
## Demand should be de-trended first
## The relationship may change over time 
fit <- lm(Demand ~ 0 + HDD, data = x)
summary(fit)
```

```{r}
df <-
  tickers_eia %>% dplyr::filter(grepl(pattern = "eia.mdist.stocks.us.dayssupply", x = tick.r)) %>%
  dplyr::mutate(key = EIAkey) %>%
  dplyr::mutate(df = purrr::pmap(list(tick.eia, key, tick.r), .f = RTL::eia2tidy)) %>%
  dplyr::select(df) %>% tidyr::unnest(df)
fig.title = "STL"
RTL::chart_zscore(
  df = df,
  title = fig.title,
  per = "yearweek",
  output = "stl",
  chart = "seasons"
)
```

```{r}
fig.title = "STL Stats"
 RTL::chart_zscore(
  df = df,
  title = fig.title,
  per = "yearweek",
  output = "stats",
  chart = "seasons"
)
```


```{r}
fig.title = "Seasonal"
 RTL::chart_zscore(
  df = df,
  title = fig.title,
  per = "yearweek",
  output = "seasonal",
  chart = "seasons"
)
```

From the STL analysis it appears that there is some consistent seasonality in the days of supply of ULSD. This seasonality appears to be quite strong and appears to coincide with months earlier in the year which is consistent with higher demand in the late winter leading to declines in days of supply. Days of supply appears to decline from January into March and then increases during the spring and summer as heating needs decline. This relationship appears to be consistent between years with the exception of 2020 as COVID-19 led to major unexpected increases in ULSD inventories. 

## Appendix-Trading Model Based on the Spread between Capacity and Stock:
```{r, eval=FALSE}

library(tidyverse)
library(purrr)
library(RTL)
library(dplyr)
library(tibble)
library(scales)

eia_df <- 
  tibble::tribble(
  ~ticker, ~param, ~type,
  "PET.WDIST1A1.W", "New England (PADD 1A )", "sink", 
  "PET.WDIST1B1.W", "Central Atlantic (PADD 1B)", "sink", 
  "PET.WDIST1C1.W", "Lower Atlantic (PADD 1C)", "sink", 
  "PET.WDISTP11.W", "East Coast (PADD 1)", "sink",
  "PET.WDISTP21.W", "Midwest (PADD 2)", "sink", 
  "PET.WDISTP31.W", "Gulf Coast (PADD 3)", "sink", 
  "PET.WDISTP41.W", "Rocky Mountain (PADD 4)", "sink", 
  "PET.WDISTP51.W", "West Coast (PADD 5)", "sink", 
  )  # create a data frame with all the tickers 
# create data pull function to access EIA API pull data and clean and convert data to the correct formation

EIAkey = "2a8946fed36919b63af3ff2f4bcaf83a"



df_stock <- eia_df %>%
  dplyr::mutate(key = EIAkey) %>%
  dplyr::mutate(data = purrr::map2(ticker, key, RTL::eia2tidy)) %>%
  dplyr::select(param, type, data) %>% unnest() %>% 
  filter(param %in% c("East Coast (PADD 1)")& (date >= "2011-01-01"))


# df_stock


fig.title = " Ending Stocks of Distillate Fuel Oil"

p1 <-
  df_stock %>% dplyr::filter(param %in% c("East Coast (PADD 1)","Midwest (PADD 2)","Gulf Coast (PADD 3)","Rocky Mountain (PADD 4)","West Coast (PADD 5)")) %>%
  ggplot(aes(x = date, y = value, color = param)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle(fig.title)

p1
# p1 <-
#   df_stock %>% dplyr::filter(param %in% c("East Coast (PADD 1)")& (date >= "2011-01-01")) %>%
#   ggplot(aes(x = date, y = value, color = param)) +
#   geom_line() +
#   scale_y_continuous(labels = scales::comma) +
#   ggtitle(fig.title)
# 
# p1

```


```{r, eval=FALSE}
library(RTL)
library(dplyr)

df_capacity <- eiaStorageCap %>% 
  filter(product == "distillates")

df_capacity

fig.title = "Capacity of Distillate Fuel Oil"
p2 <-
  df_capacity %>%
  dplyr::filter(date >= "2011-01-01")%>%
  ggplot(aes(x = date, y = value, color = product)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma)+ggtitle(fig.title)
p2


#adjust the data manually
date_vector <-  c("2021-04-02","2020-04-03","2019-04-05","2019-10-04","2018-04-06","2018-10-05","2017-04-07","2017-10-06","2016-04-01","2016-10-07","2015-04-03","2015-10-02","2014-04-04","2014-10-03","2013-04-05","2013-10-04","2012-04-06","2012-10-05","2011-04-01","2011-10-07","2010-10-01") %>% as.Date(format = "%Y-%m-%d")

df_capacity$date = date_vector


```

The stock for distillate in East Coast (PADD 1) accounts for the largest number and fluctuate seasonally. The only available data of capacity we can get from EIA API is from PADD 1, and no more information from any other PADDs. We can reasonably simplify our model by focus on the data in PADD 1 only.

```{r, eval=FALSE}
df_capacity_adj <- df_capacity %>% rename(
    c("capacity"="value")
    ) %>% 
  select(date,capacity)

df_stock_adj <- df_stock %>% rename(
    c("stock"="value")) %>% 
  select(date,stock)



df_join <- df_stock_adj %>% 
  arrange(date) %>% 
  filter(date >= "2011-01-01") %>% 
  dplyr::left_join(df_capacity_adj, by = "date") %>% 
  fill(capacity) %>% na.omit()
  
  


fig.title = "Capacity and Stock of Distillate Fuel Oil"

df_join

df_join %>% 
  pivot_longer(-date, names_to = "series",values_to = "value" ) %>% 
  ggplot(
  aes(x = date, y = value, col = series))+
  geom_line()+ggtitle(fig.title)

```
```{r, eval=FALSE}
df_spread <- df_join %>% 
  transmute(
    date = date,
    spread = capacity-stock
  )


fig.title = "Capacity and Stock of Distillate Fuel Oil"

df_spread %>% ggplot(
  aes(x = date, y = spread)
) + 
  geom_line()+ggtitle(fig.title)
```
The spread here means the difference between Capacity and Stock (capacity-stock)

```{r, eval=FALSE}
library(ggplot2)
df_HO_c1c2_Spd <- HOdf %>% 
  filter(Contracts == "HO_c1c2_Spd") %>% 
  filter(date >= "2011-01-01") %>% 
  select(date,Price)%>%
  na.omit() %>% 
  arrange(date) %>% 
  rename(
    c("HO_c1c2_Spd"="Price")
    ) 

df_HO_Apr22_Mar22_Spd<- HOdf %>% 
  filter(Contracts == "HO_Apr22_Mar22_Spd") %>% 
  filter(date >= "2011-01-01") %>% 
  select(date,Price)%>%
  na.omit() %>% 
  arrange(date) %>% 
  rename(
    c("HO_Apr22_Mar22_Spd"="Price")
    )

df_HO_Crack_Spd<- HOdf %>% 
  filter(Contracts == "HO_Crack_Spd") %>% 
  filter(date >= "2011-01-01") %>% 
  select(date,Price)%>%
  na.omit() %>% 
  arrange(date) %>% 
  rename(
    c("HO_Crack_Spd"="Price")
    )


df_join_all <-df_spread %>% 
  dplyr::left_join(df_HO_c1c2_Spd, by = "date") %>% 
  dplyr::left_join(df_HO_Apr22_Mar22_Spd, by = "date") %>% 
  dplyr::left_join(df_HO_Crack_Spd, by = "date") %>% 
  na.omit()

df_join_all$HO_c1c2_Spd = scale(df_join_all$HO_c1c2_Spd)
df_join_all$HO_Apr22_Mar22_Spd = scale(df_join_all$HO_Apr22_Mar22_Spd)
df_join_all$HO_Crack_Spd = scale(df_join_all$HO_Crack_Spd)
df_join_all$spread = scale(df_join_all$spread)


df_join_all

df_join_all_long <- df_join_all %>% pivot_longer(-date, names_to = "series",values_to = "value" ) %>% na.omit() 

p3 <- df_join_all_long%>% 
  ggplot(
  aes(x = date, y = value,col = series)
,width = 700, height = 400) + 
  geom_line()

p3
```
It looks like the Time Spread(HO_01 vs HO_02) does not have good trend with our spread. While, we do not trade on continuous spread but on the real contracts.  Then the Crack Spread and the Time Spread(Apr22 vs Mar22) would be check.
```{r, eval=FALSE}
fit1 <- stats::lm(HO_Crack_Spd ~ as.vector(spread), data = df_join_all)
summary(fit1)
```
The result above indicates that the regression between HO_Crack_Spd ~ spread is really significant since F-statistic = 231.1 and R-squared = 0.6102.

```{r, eval=FALSE}
fit2 <- stats::lm(HO_Apr22_Mar22_Spd ~ as.vector(spread), data = df_join_all)
summary(fit2)
```



The result above indicates that the regression between HO_Apr22_Mar22_Spd ~ spread is little  less significantly since F-statistic = 37.82 and R-squared = 0.2003.

As a result, the spread could be a reasonable variable to trade on crack spread based on the spread between stock and capacity.


```{r, eval=FALSE}
df_join_all %>%
  ggplot(aes(x = spread, y = HO_Crack_Spd, col = date)) +
  geom_point() +
    geom_smooth(method = "lm"
  )
```


```{r, eval=FALSE}
library(stats)

new = data.frame(spread = 0.48782800)

# utilization = as.matrix(utilization)

fit1 %>% stats::predict.lm(newdata = new,
                       interval = "prediction",
                       level = 0.99)

df_join_all
```
Here we can put the current spread value into the regression function to predict the HO_Crack_spread and compare that with the realistic level to  decide long or short the crack spread.